Hyperparameter sensitivity

### 2020-10-20 src/s_hyperparameters.Rmd
added new hyperparameter plots

m_parameters <- expand_grid(int.depth=seq(1, 7, 2),
                            n.trees=c(1, 10, 50, 100))
r_star_vals <- f_replicate_r_star(2, m_parameters, f_data_ar1)

# saveRDS(r_star_vals, "../data/ar1_sensitivity.rds")
g1 <- readRDS("../data/hypers_ar1.rds")$r_star %>% 
  melt(id.vars=c("int.depth", "n.trees")) %>% 
  group_by(int.depth, n.trees) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  mutate(default=if_else(int.depth==3&n.trees==50, 1, 0)) %>% 
  ggplot(aes(x=as.factor(n.trees), y=middle, colour=as.factor(int.depth), group=as.factor(int.depth))) +
  geom_pointrange(aes(ymin=lower, ymax=upper, shape=as.factor(default)),
                  position=position_dodge2(width=0.1)) +
  geom_line() +
  xlab("# trees") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.25)) +
  scale_y_continuous(limits=c(0.95, 1.3)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  scale_shape(guide="none") +
  ggtitle("A. GBM")
```

8 schools
```{r}
source("eight_schools.data.R")
eight_schools <- list(J=J, y=y, sigma=sigma)
model_cp <- stan_model("eight_schools_cp.stan")
model_ncp <- stan_model("eight_schools_ncp.stan")

f_data_8_schools <- function(){
  fit_ncp <- sampling(
    model_ncp, data = eight_schools,
    iter = 2000, chains = 4, refresh = 0,
    control = list(adapt_delta = 0.95)
    )
  x_ncp <- rstan::extract(fit_ncp, permuted=F)
}

m_parameters <- expand_grid(int.depth=c(1, 3, 5),
                            n.trees=c(1, 50, 200))
r_star_vals <- f_replicate_r_star(20, m_parameters, f_data_ar1)

g1a <- readRDS("../data/hypers_8_schools.rds")$r_star %>% 
  melt(id.vars=c("int.depth", "n.trees")) %>% 
  group_by(int.depth, n.trees) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  mutate(default=if_else(int.depth==3&n.trees==50, 1, 0)) %>% 
  ggplot(aes(x=as.factor(n.trees), y=middle, colour=as.factor(int.depth), group=as.factor(int.depth))) +
  geom_pointrange(aes(ymin=lower, ymax=upper, shape=as.factor(default)),
                  position=position_dodge2(width=0.1)) +
  geom_line() +
  xlab("# trees") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.7)) +
  scale_y_continuous(limits=c(0.95, 1.6)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  scale_shape(guide="none") +
  ggtitle("A. GBM")
```

@@

g1b <- readRDS("../data/hypers_cauchy.rds")$r_star %>% 
  melt(id.vars=c("int.depth", "n.trees")) %>% 
  group_by(int.depth, n.trees) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  mutate(default=if_else(int.depth==3&n.trees==50, 1, 0)) %>% 
  ggplot(aes(x=as.factor(n.trees), y=middle, colour=as.factor(int.depth), group=as.factor(int.depth))) +
  geom_pointrange(aes(ymin=lower, ymax=upper, shape=as.factor(default)),
                  position=position_dodge2(width=0.1)) +
  geom_line() +
  xlab("# trees") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.25)) +
  scale_y_continuous(limits=c(0.95, 2.8)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  scale_shape(guide="none") +
  ggtitle("A. GBM")
```

Normal
```{r}
readRDS("../data/hypers_normal_250.rds")$r_star %>% 
  melt(id.vars=c("int.depth", "n.trees")) %>% 
  group_by(int.depth, n.trees) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  mutate(default=if_else(int.depth==3&n.trees==50, 1, 0)) %>% 
  ggplot(aes(x=as.factor(n.trees), y=middle, colour=as.factor(int.depth), group=as.factor(int.depth))) +
  geom_pointrange(aes(ymin=lower, ymax=upper, shape=as.factor(default)),
                  position=position_dodge2(width=0.1)) +
  geom_line() +
  xlab("# trees") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.25)) +
  scale_y_continuous(limits=c(0.95, 2.8)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  scale_shape(guide="none") +
  ggtitle("A. GBM")
```

# RF hyperparameters
AR1
```{r}
g2 <- readRDS("../data/hypers_rf_ar1.rds")$r_star %>%
  melt(id.vars="mtry") %>%
  group_by(mtry) %>%
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>%
  mutate(default=if_else(mtry==1, 1, 0)) %>% 
  ggplot(aes(x=mtry, y=middle)) +
  geom_pointrange(aes(ymin=lower, ymax=upper, shape=as.factor(default)),
                  position=position_dodge2(width=0.1)) +
  geom_line() +
  xlab(TeX("$m_{try}$")) +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.25)) +
  scale_y_continuous(limits=c(0.95, 1.3)) +
  scale_x_continuous(breaks=c(1, 2), limits = c(0.5, 2.5)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  scale_shape(guide="none") +
  ggtitle("B. RF")
pdf("../output/hypers_ar1.pdf", width = 12, height = 6)
multiplot(g1, g2, cols=2)
dev.off()
```

8 schools
```{r}
g2a <- readRDS("../data/hypers_rf_8_schools.rds")$r_star %>%
  melt(id.vars="mtry") %>%
  group_by(mtry) %>%
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>%
  mutate(default=if_else(mtry==4, 1, 0)) %>% 
  ggplot(aes(x=as.factor(mtry), y=middle)) +
  geom_pointrange(aes(ymin=lower, ymax=upper, shape=as.factor(default)),
                  position=position_dodge2(width=0.1)) +
  geom_line(aes(group=1)) +
  xlab("# features") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.25)) +
  scale_y_continuous(limits=c(0.95, 1.6)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  scale_shape(guide="none") +
  ggtitle("B. RF")

pdf("../output/hypers_8_schools.pdf", width = 12, height = 6)
multiplot(g1a, g2a, cols=2)
dev.off()
```
Cauchy
```{r}
g2b <- readRDS("../data/hypers_rf_cauchy.rds")$r_star %>%
  melt(id.vars="mtry") %>%
  group_by(mtry) %>%
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>%
  mutate(default=if_else(mtry==12, 1, 0)) %>% 
  ggplot(aes(x=as.factor(mtry), y=middle)) +
  geom_pointrange(aes(ymin=lower, ymax=upper, shape=as.factor(default)),
                  position=position_dodge2(width=0.1)) +
  geom_line(aes(group=1)) +
  xlab("# features") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.25)) +
  scale_y_continuous(limits=c(0.95, 2.8)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  scale_shape(guide="none") +
  ggtitle("B. RF")
pdf("../output/hypers_cauchy.pdf", width = 12, height = 6)
multiplot(g1b, g2b, cols=2)
dev.off()
```

Normal
```{r}
readRDS("../data/hypers_rf_normal_250.rds")$r_star %>%
  melt(id.vars="mtry") %>%
  group_by(mtry) %>%
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>%
  ggplot(aes(x=as.factor(mtry), y=middle, group=1)) +
  geom_pointrange(aes(ymin=lower, ymax=upper), position=position_dodge2(width=0.1)) +
  geom_line() +
  xlab("# features") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.25)) +
  scale_y_continuous(limits=c(0.95, NA)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T))
```

>>>>

# m_parameters <- expand_grid(int.depth=seq(1, 7, 2),
#                             n.trees=c(1, 10, 50, 100))
# r_star_vals <- f_replicate_r_star(2, m_parameters, f_data_ar1)

# saveRDS(r_star_vals, "../data/ar1_sensitivity.rds")
g1 <- readRDS("../data/hypers_ar1.rds")$r_star %>% 
  melt(id.vars=c("int.depth", "n.trees")) %>% 
  group_by(int.depth, n.trees) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  mutate(default=if_else(int.depth==3&n.trees==50, 1, 0)) %>% 
  ggplot(aes(x=as.factor(n.trees), y=middle, colour=as.factor(int.depth), group=as.factor(int.depth))) +
  geom_pointrange(aes(ymin=lower, ymax=upper, shape=as.factor(default)),
                  position=position_dodge2(width=0.1)) +
  geom_line(position=position_dodge2(width=0.1)) +
  xlab("# trees") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.25)) +
  scale_y_continuous(limits=c(0.95, 1.3)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  scale_shape(guide="none") +
  ggtitle("A. GBM")
```

8 schools
```{r}
# source("eight_schools.data.R")
# eight_schools <- list(J=J, y=y, sigma=sigma)
# model_cp <- stan_model("eight_schools_cp.stan")
# model_ncp <- stan_model("eight_schools_ncp.stan")
# 
# f_data_8_schools <- function(){
#   fit_ncp <- sampling(
#     model_ncp, data = eight_schools,
#     iter = 2000, chains = 4, refresh = 0,
#     control = list(adapt_delta = 0.95)
#     )
#   x_ncp <- rstan::extract(fit_ncp, permuted=F)
# }
# 
# m_parameters <- expand_grid(int.depth=c(1, 3, 5),
#                             n.trees=c(1, 50, 200))
# r_star_vals <- f_replicate_r_star(20, m_parameters, f_data_ar1)

g1a <- readRDS("../data/hypers_8_schools.rds")$r_star %>% 
  melt(id.vars=c("int.depth", "n.trees")) %>% 
  group_by(int.depth, n.trees) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  mutate(default=if_else(int.depth==3&n.trees==50, 1, 0)) %>% 
  ggplot(aes(x=as.factor(n.trees), y=middle, colour=as.factor(int.depth), group=as.factor(int.depth))) +
  geom_pointrange(aes(ymin=lower, ymax=upper, shape=as.factor(default)),
                  position=position_dodge2(width=0.1)) +
  geom_line(position=position_dodge2(width=0.1)) +
  xlab("# trees") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.7)) +
  scale_y_continuous(limits=c(0.95, 1.6)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  scale_shape(guide="none") +
  ggtitle("A. GBM")
```

@@

g1b <- readRDS("../data/hypers_cauchy.rds")$r_star %>% 
  melt(id.vars=c("int.depth", "n.trees")) %>% 
  group_by(int.depth, n.trees) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  mutate(default=if_else(int.depth==3&n.trees==50, 1, 0)) %>% 
  ggplot(aes(x=as.factor(n.trees), y=middle, colour=as.factor(int.depth), group=as.factor(int.depth))) +
  geom_pointrange(aes(ymin=lower, ymax=upper, shape=as.factor(default)),
                  position=position_dodge2(width=0.1)) +
  geom_line(position=position_dodge2(width=0.1)) +
  xlab("# trees") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.25)) +
  scale_y_continuous(limits=c(0.95, 2.8)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  scale_shape(guide="none") +
  ggtitle("A. GBM")
```

Normal
```{r}
g1c <- readRDS("../data/hypers_normal_250.rds")$r_star %>% 
  melt(id.vars=c("int.depth", "n.trees")) %>% 
  group_by(int.depth, n.trees) %>% 
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>% 
  mutate(default=if_else(int.depth==3&n.trees==50, 1, 0)) %>% 
  ggplot(aes(x=as.factor(n.trees), y=middle, colour=as.factor(int.depth), group=as.factor(int.depth))) +
  geom_pointrange(aes(ymin=lower, ymax=upper, shape=as.factor(default)),
                  position=position_dodge2(width=0.1)) +
  geom_line(position=position_dodge2(width=0.1)) +
  xlab("# trees") +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.25)) +
  scale_y_continuous(limits=c(0.95, 2.8)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  scale_shape(guide="none") +
  ggtitle("A. GBM")
```

# RF hyperparameters
AR1
```{r}
g2 <- readRDS("../data/hypers_rf_ar1.rds")$r_star %>%
  melt(id.vars="mtry") %>%
  group_by(mtry) %>%
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>%
  mutate(default=if_else(mtry==1, 1, 0)) %>% 
  ggplot(aes(x=mtry, y=middle)) +
  geom_pointrange(aes(ymin=lower, ymax=upper, shape=as.factor(default)),
                  position=position_dodge2(width=0.1)) +
  geom_line(position=position_dodge2(width=0.1)) +
  xlab(TeX("$m_{try}$")) +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.25)) +
  scale_y_continuous(limits=c(0.95, 1.3)) +
  scale_x_continuous(breaks=c(1, 2), limits = c(0.5, 2.5)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  scale_shape(guide="none") +
  ggtitle("B. RF")
pdf("../output/hypers_ar1.pdf", width = 12, height = 6)
multiplot(g1, g2, cols=2)
dev.off()
```

8 schools
```{r}
g2a <- readRDS("../data/hypers_rf_8_schools.rds")$r_star %>%
  melt(id.vars="mtry") %>%
  group_by(mtry) %>%
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>%
  mutate(default=if_else(mtry==4, 1, 0)) %>% 
  ggplot(aes(x=as.factor(mtry), y=middle)) +
  geom_pointrange(aes(ymin=lower, ymax=upper, shape=as.factor(default)),
                  position=position_dodge2(width=0.1)) +
  geom_line(aes(group=1), position=position_dodge2(width=0.1)) +
  xlab(TeX("$m_{try}$")) +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.25)) +
  scale_y_continuous(limits=c(0.95, 1.6)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  scale_shape(guide="none") +
  ggtitle("B. RF")

pdf("../output/hypers_8_schools.pdf", width = 12, height = 6)
multiplot(g1a, g2a, cols=2)
dev.off()
```
Cauchy
```{r}
g2b <- readRDS("../data/hypers_rf_cauchy.rds")$r_star %>%
  melt(id.vars="mtry") %>%
  group_by(mtry) %>%
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>%
  mutate(default=if_else(mtry==12, 1, 0)) %>% 
  ggplot(aes(x=as.factor(mtry), y=middle)) +
  geom_pointrange(aes(ymin=lower, ymax=upper, shape=as.factor(default)),
                  position=position_dodge2(width=0.1)) +
  geom_line(aes(group=1), position=position_dodge2(width=0.1)) +
  xlab(TeX("$m_{try}$")) +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.25)) +
  scale_y_continuous(limits=c(0.95, 2.8)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  scale_shape(guide="none") +
  ggtitle("B. RF")
pdf("../output/hypers_cauchy.pdf", width = 12, height = 6)
multiplot(g1b, g2b, cols=2)
dev.off()
```

Normal
```{r}
g2c <- readRDS("../data/hypers_rf_normal_250.rds")$r_star %>%
  melt(id.vars="mtry") %>%
  group_by(mtry) %>%
  summarise(lower=quantile(value, 0.25),
            middle=quantile(value, 0.5),
            upper=quantile(value, 0.75)) %>%
  ggplot(aes(x=as.factor(mtry), y=middle, group=1)) +
  geom_pointrange(aes(ymin=lower, ymax=upper), position=position_dodge2(width=0.1)) +
  geom_line(group=1, position=position_dodge2(width=0.1)) +
  xlab(TeX("$m_{try}$")) +
  ylab("R*") +
  theme(text = element_text(size=16, colour="black"),
        axis.text = element_text(colour="black"),
        legend.position = c(0.7, 0.25)) +
  scale_y_continuous(limits=c(0.95, NA)) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_brewer("Int. depth", palette = "Dark2") +
  guides(colour = guide_legend(nrow = 2, byrow = T)) +
  ggtitle("B. RF")
pdf("../output/hypers_normal_250.pdf", width = 12, height = 6)
multiplot(g1c, g2c, cols=2)
dev.off()
```
